---
title: "01.import_format_data"
author: "Guillaume Papuga"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(here)
library(tidyverse)
library(stringr)
library(sf) #shapefile
library(dismo)
library(rgdal)
library(raster)
library(nngeo)
```

# Pense bête - a faire
## variables
- la temporalité de l'eau, sur 1 an et 3 ans (moyenne) > argument veget eau
- le MSAVI. Dans l'idéal, il faudrait qu'on fasse un petit bilan des indices de productivité végétale (MSAVI, MSAVI2, NDVI, autre...?) pour pouvoir choisir le plus approprié (voir les comparer?) > non argument veget eau
- raster d'altitude MNT : joindre les couches + mise au bon format
- raster de distance aux masses d'eau salée permanente (mer + lagunes d'une taille > 1 ha, à définir) : faire une couche vecteur, calculer distance aux entités, remettre à la bonne ehcelle > pour l'instant non
- distance to shore and major salt-water body (calculated from a vector file) > pour l'instant non

## code
- faire les graphiques de réponse par variable (ordonnée 1-0)
- implémenter différents algo dans SDM
- lisser le code

# Introduction
* Article
Data processed in this file belongs to a project that aims at mapping the distribution of Mediterranean temporary lagoons. 

* Format
This document is used to format data. No analysis is coded here.
All data names once processed follow the form d.something (d stands for "data", and the second part must clearly refer to the type of data).

* Environmental variables (section 1.)
The aim of this section is to format all the environment variables. They are spatial layers (rasters) that require some pre-processing to perform spatial analysis

* Occurences (section 2.)
The aim of this section is to build a database of occurence to train the model. We used different strategies to sample pixels based on known occurences of mediterranean temporary lagoon 

* Pre-processing

## a. Download the base layer
```{r}
## hydroperiod
## 1. Upload from computer
source = "/media/papuga/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/Sentinel/" # folder path
tile = list.files(source) # names of each tile

# use a loop to extract each variable
for (i in tile){
  var.name = paste ("/hydroperiod/hydroperiod_", i, "_sep2018_sep2019.tif", sep = "") # name of the variable in the initial folder
  var.path.name = as.vector(paste (source, i, var.name, sep="")) # paste the right name
  hydroperiod = raster(var.path.name) # upload the raster 
  assign(paste("hydro_", i, sep = ""), hydroperiod) # assign the raster a unique name (from its tile)
}

# join the raster
hydroperiod = mosaic(hydro_T31TDH, hydro_T31TEH, hydro_T31TEJ, hydro_T31TFH, hydro_T31TFJ, hydro_T31TGH, 
                     fun = mean, tolerance=0.05)
## 2. Plot
plot(hydroperiod)
```


## b. Define the basic setting of raster layers
```{r layers setting}
## Set the basic parameters of the project
# each time you want to stack data, you have to respect the same PER : PROJECTION - EXTENT - RESOLUTION
# it's called the standard settings of the project.

# Same projection
p.proj = (CRS("+init=epsg:2154"))

# Correct the projection of the hydroperiod layer
p.proj
projection(hydroperiod) # read the projection definition (raster dataset)
hydroperiod.93 = projectRaster(hydroperiod, 
                               crs = p.proj) # change the CRS of the raster to WGS84

# Same extent
p.extent = extent(hydroperiod.93)

# Same resolution 
p.res = res(hydroperiod.93)
```


# 1. Load Mediterranean Temporary Lagoon (MTL) data
## a. Preset
```{r plot extent}
# due to the spatial extent of the project, it is difficult to map results
# define a local extent to visualize a subset of results (Frontignan, Hérault, France)
ext.fronti = extent (756670, 768435, 6257688, 6267642)
```


## b. Polygon data (presence - study area - absence)
```{r polygon data}
## Presence
# Read the polygon data (with the sf package)
# This multi-polygons has been drawn by hand in Qgis following Openstreet map image 
lag = st_read(here::here("data", "raw", "lag_digit.shp"))
lag.93 = st_transform(lag, crs = p.proj) # modify the projection of the vector layer

## Study area : draw a buffer around polygons
study.area = buffer (as(lag.93, Class = "Spatial"), # need a sp object to use the function 'buffer'
                     width = 2000, # width in meters
                     dissolve = T) # group all the polygons together
study.area.st = st_as_sf(study.area) # back to an sf object

## Absence : delete holes 
abs.area = st_difference(study.area.st, st_union(lag.93)) # DNF to dissolve multiple polygons first
```


# 2. Environmental data
The aim is to format X dataset : 
- hydroperiod (hydrological temporality based on Tour du Valat index - Sentinel 2 images)
- msavi.spring (**median** value of M-Savi2 during the vegetative period from february to june)
- altitude (based on IGN product at 75m resolution, resampled to match the grid of the hydroperiod layer)

Note : we use the **hydroperiod** layer as a basis to define the cell size.

### a. msavi
The msavi layer is built as the median value of all images from 2019-02-01 to 2019-07-01.
```{r}
## msavi.spring
## 1. Upload from external drive
source = "/media/papuga/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/Sentinel/" # folder path
tile = list.files(source) # names of each tile
# source = "/home/papuga/Documents/couches_lag.temp/raw.layers/Sentinel/"  # if done from my computer

# use a loop to extract each variable
for (i in tile){
  folder.path = as.vector(paste (source, i, "/MSAVI2_v0/", sep="")) # define the folder
  lst = list.files(path = folder.path, pattern='tif$') # extract the list of files
  df = data.frame(lst)  %>% # create a data frame from the list of rasterlayer
    rename (cplt = lst) %>% # rename the column to avoid ambiguous name
    separate(cplt, c("a","b","c","d"), sep = "_", remove = F) %>% # separate the first column (remove = F keep the first column)
    filter(b == "20m") %>% # delete all the rows that do not contain single image
    separate(d, c("date","ext"), sep = "[.]", remove = T) %>% # separate the last column to extract the date
    mutate (date = as.Date(date, format = "%Y%m%d")) %>% # turn the code into a real date with base.r
    filter (date > "2019-02-01", # and filter for images between the two boundary date
            date < "2019-07-01")
  img.spring = as.vector (df$cplt) # extract the name of the rasterlayer that correspond to spring Sentinel images
  st = stack (as.list(paste (folder.path, img.spring, sep=""))) # convert into path & create the raster stack
  r_median = calc(st, median) # take the median of the serie
  assign(paste("msavi_", i, sep = ""), r_median) # assign the raster a unique name (from its tile)
  }

# join the rasters
msavi = mosaic(msavi_T31TDH, msavi_T31TEH, msavi_T31TEJ, msavi_T31TFH, msavi_T31TFJ, msavi_T31TGH, 
                     fun = mean, tolerance=0.05)

## 2. Check and convert (if needed) the raster layer
# Same projection
crs(msavi) # read the projection definition (raster dataset)
msavi.93 = projectRaster(msavi,
                         crs = p.proj) 

# Check
extent(msavi.93) # Extent
p.extent
res(msavi.93) # Same resolution 
p.res

## 3. Plot
plot(msavi.93)
```


### b. Altitude
The altitude layer is uploaded at a resolution of 75m, and resampled at 20m to match the hydroperiod's grid.
```{r}
## altitude
## 1. Upload from external drive
source = "/media/papuga//My Passport for Mac/spatial.data/02.geographie/bd.alti/BDALTIV2/1_DONNEES_LIVRAISON_2018-01-00245/BDALTIV2_MNT_75M_ASC_LAMB93_IGN69_FRANCE/" # folder path
code.tile = c("0675_6150", "0675_6225", "0675_6300", "0675_6375", "0750_6300", "0750_6375", "0825_6300", "0825_6375", "0900_6225", "0900_6300", "0900_6375", "0975_6225", "0975_6300", "0975_6375", "1050_6375") 
tile = paste("BDALTIV2_75M_FXX_", code.tile, "_MNT_LAMB93_IGN69.asc", sep = "") # names of each tile

# use a loop to extract each variable
for (i in tile){
  var.path.name = paste (source, i, sep = "") # paste the right name
  mnt = raster(var.path.name) # upload the raster 
  crs(mnt) = CRS('+init=EPSG:2154')# assign its CRS, here Lambert 93
  assign(paste("altitude_", i, sep = ""), mnt) # assign the raster a unique name (from its tile)
}

# join the raster
altitude = mosaic(get(objects(pattern="asc$")[1]), 
                  get(objects(pattern="asc$")[2]),
                  get(objects(pattern="asc$")[3]),
                  get(objects(pattern="asc$")[4]),
                  get(objects(pattern="asc$")[5]),
                  get(objects(pattern="asc$")[6]),
                  get(objects(pattern="asc$")[7]),
                  get(objects(pattern="asc$")[8]),
                  get(objects(pattern="asc$")[9]),
                  get(objects(pattern="asc$")[10]),
                  get(objects(pattern="asc$")[11]),
                  get(objects(pattern="asc$")[12]),
                  get(objects(pattern="asc$")[13]),
                  get(objects(pattern="asc$")[14]),
                  get(objects(pattern="asc$")[15]),
                  fun = mean, tolerance=0.05)
plot(altitude)

## 2. Check and convert (if needed) the raster layer
# Same projection
crs(altitude) # read the projection definition (raster dataset)
altitude.93 = projectRaster(altitude, 
                            crs = p.proj) 

# Check
extent(altitude.93) # Extent
p.extent
res(altitude.93) # Same resolution 
p.res

## 3. Convert the raster layer
# Extent
altitude.93.c = crop(x = altitude.93, y = p.extent) 
altitude.93.c = extend(x = altitude.93.c, y = p.extent)

# Cell size
# resample the value with resample () (package raster)
altitude.93.c.adj = resample(altitude.93.c, hydroperiod.93, method = "bilinear")

## 4. Plot
plot(altitude.93.c.adj)
```


### c. Environment-data stack
```{r}
# Ecological variables
## Save each layer to the processed file so one don't have to run the script again
## Folder path has to be changed
writeRaster(hydroperiod.93, 
            paste ("/home/papuga/Documents/couches_lag.temp/processed_layers/",
                   "hydroperiod.93",".tif", sep = ""))

writeRaster(msavi.93, 
            paste ("/home/papuga/Documents/couches_lag.temp/processed_layers/",
                   "msavi.93",".tif", sep = ""))

writeRaster(altitude.93.c.adj, 
            paste ("/home/papuga/Documents/couches_lag.temp/processed_layers/",
                   "altitude.93.c.adj",".tif", sep = ""))
```


```{r open layers}
# Read layers (to avoid running the whole script when session restarts)
msavi.93 = raster("/home/papuga/Documents/couches_lag.temp/processed_layers/msavi.wgs.tif")
msavi.93 = raster("/home/papuga/Documents/couches_lag.temp/processed_layers/msavi.wgs.tif")
altitude.93.c.adj = raster("/home/papuga/Documents/couches_lag.temp/processed_layers/altitude.wgs.c.adj.tif")

## Stack all files together
## This is a way to check that everything has been done carefully and raster layers are compatible
ecov = stack (hydroperiod.93, 
              msavi.93,
              altitude.93.c.adj)
```

