---
title: "01.import_format_data"
author: "Guillaume Papuga"
date: "11/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
require(here)
library(tidyverse)
library(stringr)
library(raster)
library(rgdal)
library(sf) #shapefile
```


# Introduction
## Article
Data processed in this file belongs to a project that aims at mapping the distribution of Mediterranean temporary lagoons. 


## Format
This document is used to format data. No analysis is coded here.
All data names once processed follow the form d.something (d stands for "data", and the second part must clearly refer to the type of data).

## Environmental variables (section 1.)
The aim of this section is to format all the environment variables. They are spatial layers (rasters) that require some pre-processing to perform spatial analysis

## Occurences (section 2.)
The aim of this section is to build a database of occurence to train the model. We used different strategies to sample pixels based on known occurences of mediterranean temporary lagoon 

# 1. Environmental data
The aim is to format X dataset : 
- hydroperiod (hydrological temporality based on Tour du Valat index - Sentinel 2 images)
- msavi.may (M-Savi2 index of the may image)
- msavi.spring (average value of M-Savi2 during the vegetative period from february to june)
- mnt (altitude based on the the original MNT)
- distance to shore and major salt-water body (calculated from a vector file)

We use the **hydroperiod** layer as a basis to define the cell size.

## a. Download the base layer
```{r}
## hydroperiod
## 1. Upload from external drive
source = "/Volumes/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/" # folder path
var.name = "hydroperiod_sep2018_sep2019.tif" # name of the variable in the initial folder
var.path.name = as.vector(paste (source, var.name, sep="")) # paste the right name
hydroperiod = raster(var.path.name)
```

## b. Define the basic setting of raster layers
```{r}
## Set the basic parameters
# each time you want to stack data, you have to respect the same PER : PROJECTION - EXTENT - RESOLUTION
# it's called the standard settings of the project and it's p.

# Same projection
p.proj = "+proj=longlat +datum=WGS84"
projection(hydroperiod) # read the projection definition (raster dataset)

# Same extent
p.extent = extent(hydroperiod)

# Same resolution 
p.res = res(hydroperiod)
```

## c. Correct the projection of the hydroperiod layer
```{r}
hydroperiod.wgs = projectRaster(hydroperiod,
                                  crs = p.proj) # change the CRS of the raster to WGS84
```


## d. Upload all other data
### msavi.may
```{r}
## msavi.may
## 1. Upload from external drive
source = "/Volumes/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/MSAVI2/" # folder path
var.name = "MSAVI2_20m_T31TEJ_20190513.tif" # name of the variable in the initial folder
var.path.name = as.vector(paste (source, var.name, sep="")) # paste the right name
msavi.may = raster(var.path.name)

## 2. Check and convert (if needed) the raster layer
# Same projection
projection(msavi.may) # read the projection definition (raster dataset)
msavi.may.wgs = projectRaster(msavi.may,
                              crs = p.proj) 

# Same extent
extent(msavi.may.wgs)

# Same resolution 
res(msavi.may.wgs)

##if the cell size is not the same, you need to convert it
# a. change the pixel size with aggregate or disaggregate

# b. resample the value with resample () (package raster)

# test = projectRaster(bioclim.var, crs = p.proj) # transform the projection into a new one (raster dataset)
```

### msavi.spring



### Environment-data stack
```{r}
# Ecological variables
## Stack all files together
## This is a way to check that everything has been done carefully and raster layers are compatible

ecov = stack (hydroperiod.wgs, 
              msavi.may.wgs)
```




# 2. Load Mediterranean Temporary Lagoon (MTL) data
## a. Dataset structure
```{r}
# The matrix is named `d.lag` and must contain 6 columns
# pix.id : indefinite pixel code (dataset specific)
# had.id : identifica
# date : the date (yyyymmdd)
# x : coordinate on the x-axis (longitude in decimal degree)
# y : coordinate on the y-axis (latitude in decimal degree)
# precision : precision of the location (expressed in meter)
# source : explicit name of the source of the imported dataset

d.lag = as.data.frame(matrix (ncol = 7,nrow = 0))
colnames (d.occ) = c("code.pop", "sp.name", "date", "x", "y", "precision", "source")
```


## b. Hand pointed data (presence)
```{r}
# load raw data
## 1. Upload from external drive
source = "/Volumes/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/" # folder path
var.name = "lag.temp.shp" # name of the variable in the initial folder
var.path.name = as.vector(paste (source, var.name, sep="")) # paste the right name
psce = st_read(var.path.name)

## 2. Check and convert (if needed) the raster layer
# Same projection
projection(psce) # read the projection definition (raster dataset)
psce.wgs = st_transform(psce, p.proj)
```

This dataset contains `r nb.row` populations.

## b. Random absence data
```{r}
# load raw data
## 1. Upload from external drive
source = "/Volumes/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/" # folder path
var.name = "absence.shp" # name of the variable in the initial folder
var.path.name = as.vector(paste (source, var.name, sep="")) # paste the right name
abs = st_read(var.path.name)

## 2. Check and convert (if needed) the raster layer
# Same projection
projection(abs) # read the projection definition (raster dataset)
abs.wgs = abs
```

## c. Synthese
```{r}
# convert each spatial layer into a dataframe
tab.abs.wgs = as.data.frame (cbind (st_coordinates(abs.wgs), # coordinates of each point in the vector
                                    rep(0, nrow(abs.wgs)))) # 1-0 lagoon value
tab.psce.wgs = as.data.frame (cbind (st_coordinates(psce.wgs), 
                                    rep(1, nrow(psce.wgs))))
tab.pa = rbind (tab.abs.wgs, tab.psce.wgs) # bind the two tables
colnames (tab.pa) = c("x", "y", "lag") # change columns name
tab.pa = (tab.pa[complete.cases(tab.pa), ]) # remove NA

# transform the matrix into a spatial point file
coordinates (tab.pa) = ~x*y # the name of the x and y columns
proj4string (tab.pa) = p.proj # attributes a projection for the object, it DOES NOT reproject
spplot(tab.pa) # plot the spatial data with attributes
```


## b. Polygon data

## 3. Synthesis
```{r}
# Copy and paste each dataset on the basic matrix
d.occ = rbind(d.occ, d.occ.cbn)

# Eliminate duplicates
d.occ = unique (d.occ)

# write data in the "processed" folder
write.table (x = d.occ, here::here("data","processed","d.occ.conv.lan.txt"), sep = ";", dec = ".")
```

