---
title: "02.sdm_lagune"
author: "Guillaume Papuga"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=FALSE, message=FALSE, warning=FALSE)
```

```{r packages}
library (tidyverse)
library (ade4)
library (here)
library (stringr)
library (ade4)
library (dismo)
library (mapdata)
library (maps)
library (maptools)
library (raster)
library (rgdal)
library (rgeos)
library (sdm)
library (sp)
library (usdm) 
library (reshape2)
```



# Loading data
First of all, you need to run the rMarkdown document named "01.import_format_data.Rmd" to load all the data in the R session
- 1. ecological variables are grouped into a stacked raster named `ecov`
- 2. p-a data from GIS handmade points are grouped in a spatial point file named `tab.pa`
- 3. p-a data with complete habitat structure (to distinguish one habitat from another) are to be created


```{r}
# presence
psc = extract(ecov, lag.93, weights = TRUE, normalizeWeights = FALSE, factors = TRUE) %>%
  melt () %>%
  dcast (Var1 + L1 ~Var2, value.var = "value") %>%
  filter (weight == 1) %>% # exclude cells not totally contained in the polygon (avoid mixels)
  dplyr::select (-c(weight, Var1)) %>% # delete the column
  rename(lag = L1) %>%
  mutate (presence = 1) %>%
  arrange (lag)

# background
bg.pts = st_sample(abs.area, 100000) # reduce the sample of background points to 100 000 due to computation time
bg = as.data.frame(extract(ecov, as(bg.pts, "Spatial"))) %>%
  mutate (lag = NA) %>%
  mutate (presence = 0) %>%
  dplyr::select (lag, hydroperiod.93, msavi.93, altitude.93.c.adj, presence)

# join
mat = bind_rows(psc, bg) %>% # with bind row from dplyr (anyway, columns are identical here, rbind would work)
             drop_na(c(hydroperiod.93, msavi.93, altitude.93.c.adj)) 

# write the table
# write.csv (mat, here::here("data", "processed", "pa.mat.csv"))
```


# PART I : Data mining

## Hydroperiod
```{r data mining hydroperiod}
# reduce absences
data.plot = mat %>% 
  group_by(presence) %>% 
  sample_n(10000)

# plot
# histogram
# get the global shape of data
ggplot(psc, aes(x = hydroperiod.93)) + 
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5) +
  geom_density(alpha=0.3, bw = 10, fill="#FF6666") +
  labs(title="Hydroperiod histogram plot",x="Hydroperiod (days)", y = "Density") +
  theme_classic()

# boxplot
# detect eventual lagoons that could be considered as outliers
ggplot(psc, aes(x = reorder(as.factor(lag), hydroperiod.93, FUN = median), y = hydroperiod.93)) +
  geom_boxplot() +
  labs(title="Hydroperiod per lagoon", x="Lagoon id", y = "Hydroperiod")
#x = reorder(lag, hydroperiod.93, FUN = median)
# binary plot with smoother
# detect trend in the presence-absence relationship 
ggplot(data.plot, aes(x = hydroperiod.93 , y = presence)) +
  geom_point(size = 0.3) +
  stat_smooth (fill="grey", colour = "darkgrey") +
  labs(title="Hydroperiod binomial plot", x = "Hydroperiod (days)", y = "Lagoon") +
  theme_classic()
```

## Msavi
```{r data mining hydroperiod}
# reduce absences
data.plot = mat %>% 
  group_by(presence) %>% 
  sample_n(10000)

# plot
# histogram
ggplot(psc, aes(x = msavi.93)) + 
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5) +
  geom_density(alpha=0.3, bw = 1, fill="#FF6666") +
  labs(title="Msavi histogram plot",x="Msavi", y = "Density") +
  theme_classic()

# binary plot with smoother
ggplot(data.plot, aes(x = msavi.93 , y = presence)) +
  geom_point(size = 0.3) +
  stat_smooth (fill="grey", colour = "darkgrey") +
  labs(title="Msavi binomial plot", x = "Msavi", y = "Lagoon") +
  theme_classic()
```

## Altitude
```{r data mining hydroperiod}
# reduce absences
data.plot = mat %>% 
  group_by(presence) %>% 
  sample_n(10000)

# plot
# histogram
ggplot(psc, aes(x = altitude.93.c.adj)) + 
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5) +
  geom_density(alpha=0.3, bw = 0.1, fill="#FF6666") +
  labs(title="Alttude histogram plot",x="Altitude (m)", y = "Density") +
  theme_classic()

# binary plot with smoother
ggplot(data.plot, aes(x = altitude.93.c.adj , y = presence)) +
  geom_point(size = 0.3) +
  stat_smooth (fill="grey", colour = "darkgrey") +
  labs(title="Altitude binomial plot", x = "Altitude (m)", y = "Lagoon") +
  theme_classic()
```

## Correlation among variables
To come


# PART II : simple GLM procedure (handmade)
## 1. variable selection
```{r}
# Only three variables in the model
```

## 2. model
```{r}
# 1. Split the data set
rawdata = mat 

# for (i in 1:50){}
# training dataset
tr.lag = sample(unique(na.omit(rawdata$lag)), 
                0.7*length(unique(na.omit(rawdata$lag))), 
                replace = FALSE)
tr.p = rawdata %>%
          filter(lag == tr.lag)

ncell_tr.lag = nrow(tr.p)

tr.a = rawdata %>%
  filter(presence == 0) %>%
  sample_n(size = ncell_tr.lag)

train = bind_rows(tr.p, tr.a)

# test dataset
te.p = rawdata %>%
          filter(lag != tr.lag)

ncell_te.lag = nrow(te.p)

te.a = rawdata %>%
  filter(presence == 0) %>%
  sample_n(size = ncell_te.lag)

test = bind_rows(te.p, te.a)

# 2. Run the model
# due to the shape of the response curve we added a square term to hydroperiod and msavi
m1 = glm(presence ~ hydroperiod.93 + 
           I(hydroperiod.93) +
           msavi.93 + 
           I(msavi.93^2) +
           altitude.93.c.adj + 
           I(altitude.93.c.adj^2), 
         data = train, family = "binomial")

summary(m1) # model summary

# 3. evaluate the model
# to come (see Hijmann too for AUC)

# 4. predict over the whole map
pred.glm = predict (ecov, m1, type = "response")

## Average the map


```

```{r}
# Plot the result of the GLM over the Frontignan area
# 1. crop the variables
pred.glm.fronti
plot(pred.glm.fronti)

# add points
df.sp1 = as.data.frame(sp1)
points (df.sp1[which (df.sp1$Occurrence == 1), c(2,3)], 
        pch = 16, col = "red", cex = 0.5)
points (df.sp1[which (df.sp1$Occurrence == 0), c(2,3)], 
        pch = 16, col = "blue", cex = 0.3)
```




# PART III : General additive model
```{r}

```




# Part IV : **sdm** package procedure
```{r}

```












