---
title: "02.sdm_lagune"
author: "Guillaume Papuga"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=FALSE, message=FALSE, warning=FALSE)
```

```{r packages}
library (tidyverse)
library (ade4)
library (here)
library (stringr)
library (ade4)
library (dismo)
library (mapdata)
library (maps)
library (maptools)
library (raster)
library (rgdal)
library (rgeos)
library (sdm)
library (sp)
library (usdm) 
```

# Loading data
First of all, you need to run the rMarkdown document named "01.import_format_data.Rmd" to load all the data in the R session
- 1. ecological variables are grouped into a stacked raster named `ecov`
- 2. p-a data from GIS handmade points are grouped in a spatial point file named `tab.pa`
- 3. p-a date with complete habitat structure (to distinguish one habitat from another) are to be created


```{r load data}
```


# PART I : simple GLM procedure (handmade)

## 1. data mining
```{r data mining}
# A first glimpse into data structure
#. Extract values from the rasterstack
env = raster::extract(ecov.crop, tab.pa)# extract values from a raster (our environmental layers) at the location of specie
df = cbind(tab.pa@data, env) # attach the presence-absence data to the data
```

## 2. variable selection
```{r}

```

## 3. model
```{r}
# 1. Split the data set

# 2. Run the model
m1 <- glm(lag ~ hydroperiod_sep2018_sep2019 +
                I(hydroperiod_sep2018_sep2019^2) +
                layer + 
                I(layer^2) +
                MSAVI2_20m_T31TEJ_20190722 + 
                I(MSAVI2_20m_T31TEJ_20190722^2), 
          data = df, family = "binomial")
summary(m1)

pred.glm.c = predict (ecov.crop, m1, type = "response")
plot(pred.glm.c)

pred.glm = predict (ecov, m1, type = "response")
plot(pred.glm)

writeRaster(x = pred.glm.c, 
            filename = "/Volumes/My Passport for Mac/spatial.data/09.projets.specifiques/couches_lag.temp/raw.layers/raw.res/test.pred.glm.crop.tif")


# add points
df.sp1 = as.data.frame(sp1)
points (df.sp1[which (df.sp1$Occurrence == 1), c(2,3)], 
        pch = 16, col = "red", cex = 0.5)
points (df.sp1[which (df.sp1$Occurrence == 0), c(2,3)], 
        pch = 16, col = "blue", cex = 0.3)

# 3. evaluate the model
```









# PART II : **sdm** package procedure










